---
title: "Generating SOC and biomass data with mrorganic"
author: "Pascal Führlich"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating SOC and biomass data with mrorganic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`mrorganic` greatly simplifies downloading and preparing SOC and biomass data for further scientific work.
The prefix "mr" of `mrorganic` indicates that it is a `madrat`-based data package.
`madrat` is a package and a framework that structures and supports data preparation
with various tools e.g. aggregation or caching.


# Installation
```{r, echo = TRUE, eval = FALSE}
# enable R to discover and download R packages from PIK
options(repos = c(CRAN = "@CRAN@", pik = "https://rse.pik-potsdam.de/r/packages"))
install.packages("mrorganic")
```

# Data Package
Like most `madrat`-based packages, `mrorganic` includes a function that runs all preparation steps and creates a tgz
archive which includes all data sets. For `mrorganic` some plots are also included. Run it like this:
```{r, echo = TRUE, eval = FALSE}
# enable R to discover and download R packages from PIK
library(mrorganic)
path <- retrieveData("organic", rev = 2)
message(path, " created.")
```
All data will be on ISO country resolution. To instead aggregate to GTAP11 regions:
```{r, echo = TRUE, eval = FALSE}
# enable R to discover and download R packages from PIK
retrieveData("organic", rev = 2, regionmapping = "regionmappingGTAP11.csv")
```

# SOC Data
To get SOC data including land type data as weights:
```{r, echo = TRUE, eval = TRUE}
library(mrorganic)
soc <- calcOutput("SOCbyLandType", aggregate = FALSE, supplementary = TRUE)
plotMap(soc$x, name = "SOC")
```
We set `aggregate = FALSE` here, because by default `calcOutput` would aggregate to
the regions currently defined by `getConfig("regionmapping")`, but we want gridded data.
With `supplementary = TRUE` we get metadata in addition to the SOC data. In this case
we care about the land type data which we want to plot next:
```{r, echo = TRUE, eval = TRUE}
landShares <- round(soc$weight, 8) / dimSums(soc$weight, dim = 3) # calculate shares
getComment(landShares) <- "unit: 1"
plotMap(landShares, name = "land cover share")
```

# Biomass Data
The biomass dataset has multiple subtypes that can be selected like this:
```{r, echo = TRUE, eval = TRUE}
library(mrorganic)
biomassAbove <- calcOutput("BiomassbyLandType", subtype = "aboveground")
plotMap(biomassAbove, name = "biomass_aboveground")

biomassBelow <- calcOutput("BiomassbyLandType", subtype = "belowground")
plotMap(biomassBelow, name = "biomass_belowground")
```
